Commutative Character Sheaves

### 2015-11-24 followup.tex
Moving computers

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item compact open subgroups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that \[\,^\circ K = J^0\cdots J^d;\]
\item a continuous representation $\sigma_0$ of $J^0$;
\item a quasicharacter $\varphi_i : J^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
These representations $(\sigma_0, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties, which allow Yu to construct, for $i=0, \ldots, d-1$, further:
\begin{enumerate}
\item a quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, and the structure of a $\Fq$-vector space on $V_{i+1}$, compatible with the abelian group structure above;
\item a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$;
\item an additive character $\psi_{i+1} : C_{i+1} \to \EEx$;
\item a group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$;
\item a quotient $J^{i+1} \to H^{i+1}$, where $H^{i+1}$ is a Heisenberg $p$-group, together with an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $(V_{i+1}, \langle\ ,\ \rangle_{i+1}, C_{i+1}, \psi_{i+1})$.
\end{enumerate}
The pair $(f_i, j_{i+1})$ is a symplectic action of $J^0\cdots J^i$ on $H^{i+1}$, so it defines a group homomorphism 
\begin{equation}\label{eq:symaction}
\begin{tikzcd}
h_{i+1} : (J^0\cdots J^i)\ltimes J^{i+1} \arrow{r} & (J^0\cdots J^i)\ltimes H^{i+1} \arrow{r} & \Sp(V_{i+1})\ltimes V_{i+1}^\sharp.
\end{tikzcd}
\end{equation}
%

From this information, Yu constructs a representation $\,^\circ \rho_i$ of $J^0\cdots J^i$, for each $i=0, \ldots, d$, such that $\,^\circ \rho_d = \,^\circ \rho$. We will need some detail about this construction, so we will review it here.
%Let $\varpi_1$ be the Heisenberg representation of the Heisenberg group $V_1^\sharp$, with central character $\psi_1$.
%Let $\varrho_1$ be the Weil representation of $\Sp(V_1)$ with the same representation space as $\omega_1$.
Let $\omega_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, with the same representation space as the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
\[
\begin{tikzcd}
\  & J^{i+1} \arrow{d} & \\
Sp(V_{i+1}) \ltimes V_{i+1}^\sharp & \arrow{l}[swap]{h_1} (J^0 \cdots J^i) \ltimes J^{i+1} \arrow{r} \arrow{d} & J^0 \cdots J^i J^{i+1}\\
& (J^0 \cdots J^i) & 
\end{tikzcd}
\]
Then $h_1^*(\omega_1)$ is a representation of $J^0\ltimes J^1$, determined by the quasicharacter $\varphi_1$ of $J^1$. 
Write $\inf(\sigma_0\otimes \varphi_0)$ for the representation of $J^0 \ltimes J^1$ obtained by pulling back $\sigma_0\otimes \varphi_0$ along $J^0 \ltimes J^1 \to J^0$. 
By \cite{yu:01a}*{}, the representation 
$h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 \ltimes 1) \cap (1 \ltimes J^1)$, so $h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ descends to a representation $\,^\circ \rho_1$ of $J^0 J^1$.  
%
As above, the representation 
$h_{2}^*(\omega_{2}) \otimes \inf(\,^\circ \rho_1)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 J^1 \ltimes 1) \cap (1 \ltimes J^2)$, so $h_2^*(\omega_2) \otimes \inf(\,^\circ \rho_1)$ descends to a representation $\,^\circ \rho_2$ of $J^0 J^1 J^2$.  
Continuing in this way, one constructs the representation $\,^\circ \rho_d$ of $J^0\cdots J^d$.

The character of $\,^\circ \rho_d$ on $J^0\cdots J^d$ satisfies
\[
\trace(\,^\circ \rho_d)(k_0 \cdots k_d) = \sigma_0(k_0) \left( \varphi_0(k_0) \cdots \varphi_d(k_d) \right) ,
\]
with ...

@@

The first step in the geometrization of the type $(\,^\circ K,\,^\circ \rho)$ is to replace $\,^\circ K$ with a group scheme $G$ over $\Fq$ so that $G(\Fq) = \,^\circ K$, canonically. 
By \cite{yu:03a}*{Prop 10.2} there is a canonical affine smooth group scheme $\underline{G}$ of finite type over $\OK$ such that \[\underline{G}(\OK)= \,^\circ K\] 
For every positive integer $n$, let $G_n$ be the Greenberg transform of $\underline{G} \times_{\Spec{\OK}} \Spec{\OK/\mathfrak{m}_K^n}$, where $\mathfrak{m}_K$ is the maximal ideal in $\OK$.
Then $G_n$ is a smooth group scheme over $\Fq$ and $G_n(\Fq) = \underline{G}(\OK/\mathfrak{m}_K^n)$, canonically.
Let $G$ be the Greenberg transform of $\underline{G}$.
Then $G$ is a group scheme over $\Fq$ and $G(\Fq) = \underline{G}(\OK) = \,^\circ K$, canonically.

In order to complete the geometrization of the type $(\,^\circ K,\,^\circ \rho)$, we will need more smooth models.
Fortunately, as above, Yu has already done the heavy work.
%
As explained in \cite{yu:03a}*{\S 10.4}, the
the group scheme $\underline{G}$ comes with the following auxillary data:
\begin{enumerate}
\item[(Ji)]
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\underline{J}^i(\OK) = J^i,\]
\item[(Gj)] affine smooth group schemes $\underline{G}^{j}$, for $j=0, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = J^0\cdots J^i.
\] 
\end{enumerate}

Moreover, the quotient

Regarding this construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:03a} concludes with the following provocative observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}
%
The objective of this section is to push this observation further by geometrizing the representation $\,^\circ \rho$ of $\underline{G}$.

>>>>

Every Yu type $(\,^\circ K,\,^\circ \rho)$ may be constructed directly from the following information, derived from a Yu datum:
\begin{enumerate}
\item compact open subgroups $J^i\subset \,^\circ K$, for $i=0, \ldots d$, such that \[\,^\circ K = J^0\cdots J^d;\]
\item a continuous representation $\sigma_0$ of $J^0$;
\item a quasicharacter $\varphi_i : J^i \to \CC^\times$, for $i=0, \ldots d$.
\end{enumerate}
These representations $(\sigma_0, \varphi_0, \ldots , \varphi_d)$ enjoy certain properties, which allow Yu to construct, for $i=0, \ldots, d-1$, further:
\begin{enumerate}
\item a quotient $J^{i+1} \to V_{i+1}$ to a finite abelian group, and the structure of a $\Fq$-vector space on $V_{i+1}$, compatible with the abelian group structure above;
\item a symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$, where $C_{i+1}$ is a one-dimensional vector space over $\Fq$;
\item an additive character $\psi_{i+1} : C_{i+1} \to \EEx$;
\item a group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$;
\item a quotient $J^{i+1} \to H^{i+1}$, where $H^{i+1}$ is a Heisenberg $p$-group, together with an isomorphism $j_{i+1} : H^{i+1} \to V_{i+1}^\sharp$, where $V_{i+1}^\sharp$ is the Heisenberg group determined by $(V_{i+1}, \langle\ ,\ \rangle_{i+1}, C_{i+1}, \psi_{i+1})$.
\end{enumerate}
The pair $(f_i, j_{i+1})$ is a symplectic action of $J^0\cdots J^i$ on $H^{i+1}$ in the sense of \cite{yu:construction}, so, together with the quotient $J^{i+1} \to H^{i+1}$, it defines a group homomorphism 
\begin{equation}\label{eq:symaction}
\begin{tikzcd}
h_{i+1} : (J^0\cdots J^i)\ltimes J^{i+1} \arrow{r} 
%& (J^0\cdots J^i)\ltimes H^{i+1} \arrow{r} 
& \Sp(V_{i+1})\ltimes V_{i+1}^\sharp.
\end{tikzcd}
\end{equation}
%

From this information, Yu constructs a representation $\,^\circ \rho_i$ of $J^0\cdots J^i$, for each $i=0, \ldots, d$, such that $\,^\circ \rho_d = \,^\circ \rho$. We will need some detail about this construction, so we will review it here.
%Let $\varpi_1$ be the Heisenberg representation of the Heisenberg group $V_1^\sharp$, with central character $\psi_1$.
%Let $\varrho_1$ be the Weil representation of $\Sp(V_1)$ with the same representation space as $\omega_1$.
Let $\omega_{i+1}$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V_{i+1})\ltimes V_{i+1}^\sharp$, with the same representation space as the Heisenberg representation of the Heisenberg group $V_{i+1}^\sharp$.
\[
\begin{tikzcd}
\  & J^{i+1} \arrow{d} & \\
Sp(V_{i+1}) \ltimes V_{i+1}^\sharp & \arrow{l}[swap]{h_1} (J^0 \cdots J^i) \ltimes J^{i+1} \arrow{r} \arrow{d} & J^0 \cdots J^i J^{i+1}\\
& (J^0 \cdots J^i) & 
\end{tikzcd}
\]

Now $h_1^*(\omega_1)$ is a representation of $J^0\ltimes J^1$, determined by the quasicharacter $\varphi_1$ of $J^1$. 
Write $\inf(\sigma_0\otimes \varphi_0)$ for the representation of $J^0 \ltimes J^1$ obtained by pulling back $\sigma_0\otimes \varphi_0$ along $J^0 \ltimes J^1 \to J^0$. 
By \cite{yu:01a}*{}, the representation 
$h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 \ltimes 1) \cap (1 \ltimes J^1)$, so $h_1^*(\omega_1) \otimes \inf(\sigma_0\otimes \varphi_0)$ descends to a representation $\,^\circ \rho_1$ of $J^0 J^1$.

Likewise, the representation 
$h_{2}^*(\omega_{2}) \otimes \inf(\,^\circ \rho_1)$ of $J^0 \ltimes J^1$ is trivial on $(J^0 J^1 \ltimes 1) \cap (1 \ltimes J^2)$, so $h_2^*(\omega_2) \otimes \inf(\,^\circ \rho_1)$ descends to a representation $\,^\circ \rho_2$ of $J^0 J^1 J^2$.

Continuing in this way, one constructs the representation $\,^\circ \rho_d$ of $J^0\cdots J^d$. 
%
The character of $\,^\circ \rho_d$ on $\,^\circ K = J^0\cdots J^d$ satisfies
\[
\trace(\,^\circ \rho_d)(k_0 \cdots k_d) = \sigma_0(k_0) \left( \varphi_0(k_0) \cdots \varphi_d(k_d) \right) ,
\]
for $k_i\in J^i$.

@@

The first step in the geometrization of the type $(\,^\circ K,\,^\circ \rho)$ is to replace $\,^\circ K$ with a group scheme $G$ over $\Fq$ so that $G(\Fq) = \,^\circ K$, canonically. 
By \cite{yu:models}*{Prop 10.2} there is a canonical affine smooth group scheme $\underline{G}$ of finite type over $\OK$ such that \[ 
\underline{G}(\OK)= \,^\circ K.
\] 
Set $S = \Spec{\OK}$ and $S_n = \Spec{\OK/\mathfrak{m}^n}$, where $\mathfrak{m}_K$ is the maximal ideal in $\OK$, and $n$ is a positive integer $n$.
Let $G_n$ be the Greenberg transform of $\underline{G} \times_{S} S_n$.
Then $G_n$ is a smooth group scheme over $\Fq$ and 
\[
G_n(\Fq) = \underline{G}(\OK/\mathfrak{m}_K^n),
\]
canonically.
%
Let $G$ be the Greenberg transform of $\underline{G}$.
Then $G$ is a group scheme over $\Fq$ and 
\[
G(\Fq) = \,^\circ K,
\]
canonically.

%
As explained in \cite{yu:03a}*{\S 10.4}, the
the group scheme $\underline{G}$ comes with the following auxillary data:
\begin{enumerate}
\item[(Ji)]
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\underline{J}^i(\OK) = J^i,\]
\item[(Gj)] affine smooth group schemes $\underline{G}^{j}$, for $j=0, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = J^0\cdots J^i.
\] 
\end{enumerate}

\newcommand{\reductive}{{\operatorname{red}}}

With these group schemes, the other groups appearing in Section~\ref{ssec:review} may also be put into the framework needed for geoemtrization.

Write $\underline{J}^{i+1}_\Fq$ for the special fibre $\underline{J}^{i+1}\times_S S_1$ of $\underline{J}^{i+1}$. The Heisenberg p-group $H^{i+1}$ group may be realized as $H^{i+1} = H^{i+1}_{\Fq}(\Fq)$, where $H^{i+1}_{\Fq}$ a quotient of $\underline{J}^{i+1}_{\Fq}$ as an algebraic group. Then the quotient $J^{i+1}\to H^{i+1}$ is realized as the composition 
\[
\underline{J}^{i+1}(\OK) \to \underline{J}^{i+1}(\Fq) = \underline{J}^{i+1}_\Fq(\Fq) \to H^{i+1}_{\Fq}(\Fq).
\]

The quotient  $J^{i+1} \to V_{i+1}$ admits a similar description, realizing $V_{i+1}$ as a quotient of the special fibre of $\underline{J}^{i+1}$

 and $\underline{J}^{i+1, \reductive}_\Fq$ for the reductive quotient of $\underline{J}^{i+1}_\Fq$.

In order to complete the geometrization of the type $(\,^\circ K,\,^\circ \rho)$, we will need more smooth models.
Fortunately, as above, Yu has already done the heavy work.

Regarding this construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation.

\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}
%
The objective of this section is to push this observation further by geometrizing the representation $\,^\circ \rho$ of $\underline{G}$.
